---
title: "Experiment 0 pilot"
author: "Matt Green"
output:
  html_document:
    toc: yes
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---  

This document aims to be a complete and up to date record of the analysis for vagueness 0, the pilot experiment.

Knitr doc options.

```{r docOpts}
library(knitr)
knit_hooks$set(crop=hook_pdfcrop)
```

Load libraries.

```{r loadLibraries}
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
```

Source additional R files with functions for computing confidence intervals.

```{r sourceCIFunctions}
source("CIfunctions.R")
```

# Preprocessing

Get the data. 

```{r getData}
if( file.exists('data.txt') ) 
	{
	d <- read.table("data.txt", header=FALSE) 
} else {
	system("cat ../experimentCode/output/*.dat > data.txt", wait=TRUE)
}
mynames <- list("gap", "target", "side", "vagueness", "pair", "Version", "Left",
								"Right", "Text", "Condition", "correctAnswer", "subject",
								"actualAnswer", "isCorrect", "rt")
names(d) <- mynames
d$obs <- 1:nrow(d)
d$trial <- 0
for (s in unique(d$subject)){
  d[d$subject==s, "trial"] <- 1:nrow(subset(d,subject==s))
}
# For some reason there is one rt of -6. Set it to NA
d[d$rt==-6,"rt"] <- NA
# make a df
d <- data.frame(
  obs = d$obs, 
  trial = d$trial, 
  subject = d$subject, 
  item = factor(d$pair, label = c("2:4", "3:5", "6:8", "7:9", "2:6", "7:3", "8:4", "5:9")),
  gap = factor(d$gap, levels = c("BigGap", "SmallGap"), labels = c("Big Gap", "Small Gap")), 
  quan = factor(d$target, levels = c("BigTarget", "SmallTarget"), labels = c("Big Target", "Small Target")), 
  vag = factor(d$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")), 
  side = d$side, 
  err = 1 - d$isCorrect, 
  rt = d$rt)
d$subit <- "Not subitizable"
d$subit[d$item == "2:4"] <- "Subitizable"
d$subit[d$item == "3:5"] <- "Subitizable"
d$subit[d$item == "2:6"] <- "Subitizable"
d$subit[d$item == "7:3"] <- "Subitizable"
d$subit = as.factor(d$subit)
d <- subset(d, select = c("obs", "trial", "subject", "item", "gap", "quan", 
                          "vag", "subit", "side", "err", "rt"))
```

Size of the data is \Sexpr{nrow(d)} trials. 1 was removed for being minus 6 rt.
Remove erroneous response trials. There are \Sexpr{nrow(subset(d,err==1))} error trials.

```{r removeErrorTrials}
# by setting the RT to NA
d[d$err==1,"rt"] <- NA
```

Do some factors and contrasts.  Set subject and item as factors. Sum code vagueness and subitizability.

```{r postProcessData}
d$subject <- as.factor(d$subject)
d$item <- as.factor(d$item)
d$vag <- relevel(d$vag, ref="Vague")
contrasts(d$vag) <- contr.sum(2)/2
d$subit <- relevel(d$subit, ref="Subitizable")
contrasts(d$subit) <- contr.sum(2)/2
contrasts(d$quan) <- contr.sum(2)/2
contrasts(d$gap) <- contr.sum(2)/2
```

Check the contrasts.

```{r checkContrasts}
contrasts(d$vag)
contrasts(d$subit)
contrasts(d$quan)
contrasts(d$gap)
```

Add transformations of rt

```{r addTransform}
d$logRT <- log(d$rt)
d$recipRT <- -10000/d$rt
```

For the untrimmed data, show effects of transforming rt.

```{r plotDistributions, fig.width=4, fig.height=2}
xs = melt(d,  
					measure.vars = c("rt","logRT","recipRT"))
densityplot(
	~value|variable,
	data=xs,
	xlab=NULL,
	yaxt=F,
  scales=list(x=list(relation="free", rot=45), 
   						y=list(relation="free", at=NULL)),
	cex=.2,
	plot.points=FALSE,
	layout=c(3,1)
	)
```

Trim the data. This function doesn't work well if there are NA values in the response, so explicitly remove them first.

```{r trimData}
dummy <- subset(d, !is.na(rt))
data.trim <- perSubjectTrim.fnc(dummy, response="rt",subject="subject", trim=2.5)$data
```

Check the distribution of the trimmed data to see if a transformation is suggested.

```{r checkTrim, fig.width=4, fig.height=2}
xs = melt(data.trim,  
					measure.vars = c("rt","logRT","recipRT"))
densityplot(
	~value|variable,
	data=xs,
	xlab=NULL,
	yaxt=F,
  scales=list(x=list(relation="free", rot=45), 
   						y=list(relation="free", at=NULL)),
	cex=.2,
	plot.points=FALSE,
	layout=c(3,1)
	)
```

Select either trimmed or untrimmed data by comparing this plot with the untrimmed data. Also use this plot to choose the most normal data (logRT).

```{r selectTrim}
dat <- data.trim
```

# RT modelling

Do a first lmer model. You can't put subit or gap over items because they are between-items.

```{r, lmers}
lmer2 <- lmer(data=dat, logRT~subit*vag + (subit*vag|subject) + (vag|item))
# lmer3 is closer to the original anova than lmer2
lmer3 <- lmer(data=dat, logRT~subit*vag+quan+gap + (subit*vag|subject) + (vag|item))
```

```{r, summaryLMER, results='asis'}
print(xtable(summary(lmer3)$coefficients, digits=3), type="html")
```

## subitizable

```{r, subitRT, results='asis'}
print(xtable(summary(lmer(data=dat, subset=subit=="Subitizable",logRT~vag+quan+gap +(vag|subject)+(vag|item)))$coefficients, digits=3), type='html')
```

## not subitizable

```{r, notsubitrT, results='asis'}
print(xtable(summary(lmer(data=dat, subset=subit=="Not subitizable", logRT~vag+quan+gap +(vag|subject)+(vag|item)))$coefficients, digits=3), type='html')
```

# Error rate analysis

```{r errorrates, results='asis'}
print(xtable(summary(
	glmer(data=d, err~vag*subit+quan+gap + (vag|subject) + (1|item), 
				family=binomial, 
				control = glmerControl(optimizer = "bobyqa")
				)
	)$coefficients,
	digits=3), 
	type='html')
```

# Plots of significant effects

```{r plot_sig, fig.width=7.5, fig.height=3}
a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","subit","item"), idvar="subject")
as <- subset(a,subit=="Subitizable")
as$item <- as.character(as$item)
as$item[as$item=="7:3"] <- "3:7"
as$item <- factor(as$item)
pd <- position_dodge(.05)
plot1 = ggplot(data=as, aes(y=logRT, x=item, ymin=logRT-ci, ymax=ci+logRT, fill=vag, group=vag)) + 
  geom_line(position=pd) +
  geom_errorbar(width=.2, position=pd)+
  geom_point(pch=21, size=4, position=pd) +
  scale_fill_grey(name=element_blank(),start=0,end=1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key=element_blank()) +
  ggtitle("Response time")

a <- summarySEwithin(data=d, measurevar="err", withinvars=c("vag","subit","item"), idvar="subject")
as <- subset(a,subit=="Subitizable")
as$item <- as.character(as$item)
as$item[as$item=="7:3"] <- "3:7"
as$item <- factor(as$item)
pd <- position_dodge(.05)
plot2 = ggplot(data=as, aes(y=err, x=item, ymin=err-ci, ymax=ci+err, group=vag, fill=vag))+
  geom_line(position=pd) +
  geom_errorbar(width=.2, position=pd)+
  geom_point(pch=21, size=4, position=pd) +
  scale_fill_grey(name=element_blank(),start=0,end=1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key=element_blank()) +
  ggtitle("Error rate")

grid.arrange(plot1, plot2, nrow=1, top="Subitizable only")
```

```{r barcharts, fig.width=9, fig.height=3}
bwtheme <- standard.theme("pdf", color=FALSE)
b1=barchart(data=ddply(dat,.(vag,subit),summarise, logRT=mean(logRT)), 
						logRT~subit, 
						group=vag, auto.key=list(space='right', cex=.7), par.settings=bwtheme)
b2=barchart(data=ddply(d,.(vag,subit),summarise, Error_Rate=mean(err)), 
						Error_Rate~subit, 
						group=vag, auto.key=list(space='right', cex=.7), par.settings=bwtheme)
grid.arrange(b1,b2,ncol=2)
```

Subitizability x Vagueness x Quantity

```{r plotMeansa, fig.width=7, fig.height=3}
a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","quan","subit"), idvar="subject")
pd <- position_dodge(.1)
	ggplot(
		a, 
		aes(y=logRT, x=vag, group=quan, shape=quan, col=quan, ymin=logRT-ci, ymax=ci+logRT)
	) + 
	geom_point(size=4, position=pd) + 
	geom_line(position=pd) + 
	facet_grid(~subit) + 
	geom_errorbar(width=.1, position=pd) + 
	theme_bw() + 
	theme(panel.grid.major=element_blank()) 
```

```{r qqplotLMER, fig.width=3, fig.height=3.5}
qqPlot(residuals(lmer2),cex=.5)
```
