---
title: "Experiment 3 of 4"
author: "Matt Green"
output:
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    code_folding: show
    theme: default
---  
<style>
table, th, td {
    border: 1px grey;
    border-collapse: collapse;
    padding:2px 5px 2px 5px
}  
</style>

# Experiment 3

```{r loadLibraries, message=FALSE}
library(knitr); library(MASS); library(ggplot2); library(lme4); library(xtable); library(plyr); library(lattice); library(LMERConvenienceFunctions); library(car); library(grid); library(lmerTest); library(reshape2); library(ggplot2); library(languageR); library(pander); source('summarySEwithin2.R'); library(Hmisc)
```

Experiment Structure: 
Each subject saw 192 trials, being 3 cycles of 64 trials. There were 38 participants. There were 7296 rows in the full data set.

Trial Procedure: 
First an instruction that was a referring expression was presented until participants dismissed it with a key press. Then the squares with dots were presented until the participant responded with a key press, with a 60 second timeout if they did not respond.

```{r showInstructions, echo=FALSE, cache=TRUE}
v = c("crisp", "crisp", "vague", "vague")
n = c('number', 'number', 'number', 'number')
t = c("matching", "comparison", "matching", "comparison")
e = c("Choose a square with 6 dots", "Choose a square with fewer than 20 dots", "Choose a square with about 10 dots", "Choose a square with far fewer than 20 dots")
i = data.frame(v, n, t, e)
names(i) = c("vagueness", "instruction format", "selection task", "example instruction")
pander(i, justify=c('center', 'left', 'left', 'left'))
```

```{r showItems, echo=FALSE, cache=TRUE}
items = c(1,2,3,4,5,6,7,8)
i = data.frame(Item = c("1","2","3","4"),
               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
names(i) = c('Item', 'Numbers of dots')
pander(i, justify=c('center', 'center'))
```

```{r getData, echo=T}
# data.txt
rawdata <- read.table('preprocessing/data_processed.txt', sep=' ')
```

```{r trimat25sd}
trimdata <- perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data
rtdata <- trimdata
```

```{r rtplot1, fig.width=7, fig.height=3}
# this plot uses a precomputed summary done by summarySEwithin
# and data trimmed at 2.5 sd per subject
rtdataplot1 <- summarySEwithin2(data=perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data,
                                measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
dodge = position_dodge(width=0.2)
rtdataplot1$Condition <- as.factor(paste(sep=' ', rtdataplot1$Selection, rtdataplot1$Vagueness))
rtplot=ggplot(rtdataplot1,
              aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci,
                group=Condition, shape=Condition, fill=Condition)) +
  geom_line(position=dodge) +
  geom_errorbar(width=.2, position=dodge) +
  geom_point(size=2, position=dodge) +
  scale_shape_manual("",values = c(22, 22, 21, 21)) +
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Response time") +
  ylab("Respone time (log (ms))") + 
  xlab("") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), aspect.ratio=1,
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.background=element_rect(fill=NA, color='white')) +
  facet_grid(~Selection) 
rtplot
```

```{r rtplotHmisc, fig.width=7, fig.height=3}
# n.b. rtdata has been trimmed at 2.5 sd: use rawdata if you want untrimmed data
rawdata$Condition <- as.factor(paste(sep=' ', rawdata$Selection, rawdata$Vagueness))
ggplot(rawdata, aes(y=RT_log, x=Item, group=Condition, shape=Condition, fill=Condition)) + 
  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(0.2)) +
  scale_shape_manual("",values = c(22, 22, 21, 21)) +
  scale_fill_manual("",values=c("black","white","black","white")) +
  facet_grid(~Selection) +
  theme(
    panel.background=element_rect(fill=NA, color='black'),
    panel.grid.major=element_line(color='grey95'), 
        #panel.grid.minor = element_blank(),
        legend.key=element_blank(), 
        aspect.ratio=1,
        axis.text.x=element_text(angle = 45, hjust = 1)
        )
```


```{r rtlmer, cache=TRUE}
rtlmer = lmerTest::lmer(data=rtdata,
                        RT_log ~ c_Vag * c_Sel + c_Itm +
                          (1 + c_Vag * c_Sel + c_Itm | Subject))
```

```{r rtlmersummaryxtable, results='asis', cache=TRUE, echo=F}
print(xtable(summary(rtlmer)$coef, digits=c(0,2,2,1,1,4)),type='html')
```

```{r seplmersforcomparison, results='asis', cache=TRUE, echo=-3}
s1 <- subset(rtdata, subset=Selection=='Comparison')
s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
```

```{r seplmersformatching, results='asis', cache=TRUE, echo=-3}
s2 <- subset(rtdata, subset=Selection=='Matching')
s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
```
